---
title: "Waterbody Connectivity across Watersheds"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
---

This document assesses the connectivity and flow direction of streams and rivers that flow from the watershed 08N to the target watersheds 08P, 08M, 08L and 08K.

## Results {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, fig.width = 10, fig.height = 6)
library(tidyverse)
library(sf)
library(leaflet)
library(DT)
library(terra)
library(bcdata)
```

### Study Area

```{r load_data}
# BC boundary polygon
bc = bcmaps::bc_bound() |> dplyr::summarise()

# Watershed drainages
wscd = bcmaps::wsc_drainages()

# Merge by SUB_DRAINAGE_AREA_CD
wscd_s = wscd |> 
  sf::st_zm() |> 
  dplyr::group_by(SUB_DRAINAGE_AREA_CD) |> 
  dplyr::summarise()

# Target watershed codes.
wscd_t = wscd_s |> 
  dplyr::filter(SUB_DRAINAGE_AREA_CD %in% c("08N","08P","08M","08L","08K"))

# Trim to BC
wscd_t = sf::st_intersection(wscd_t, bc)

em_lake = bcdc_query_geodata('freshwater-atlas-lakes') |> 
  filter(GNIS_NAME_1 == 'Emerald Lake') |> 
  filter(WATERSHED_GROUP_ID == 79) |> 
  collect() |> 
  sf::st_zm()

em_lake_c = em_lake |> 
  sf::st_centroid()
  
base_wd_map = ggplot() + 
  geom_sf(data = wscd_t, aes(fill = SUB_DRAINAGE_AREA_CD)) + 
  geom_sf_text(data = wscd_t, aes(label = SUB_DRAINAGE_AREA_CD)) +
  geom_sf(data = em_lake_c, color = 'red') +
  ggrepel::geom_label_repel(
    data = em_lake,
    label = 'WD Detection in \nYoho National Park',
    aes(geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
    nudge_y = 100000,
    nudge_x = 50000
  ) +
  labs(fill = 'Watershed Code', title = 'Major Watersheds in SE British Columbia') +
  ggthemes::theme_map() + 
  ggspatial::annotation_scale(location = 'br')

base_wd_map

# insp = read_sf('W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years.gpkg')
```

```{r border_of_08N_and_target_watersheds}

### Border of 08N that Touches Target Watersheds

border_08N = wscd_t |> 
  dplyr::filter(SUB_DRAINAGE_AREA_CD == '08N') |> 
  st_cast("LINESTRING") |> 
  sf::st_intersection(wscd_t[wscd_t$SUB_DRAINAGE_AREA_CD != '08N',]) |> 
  dplyr::summarise()

# Simplify geometry
border_08N_s = rmapshaper::ms_simplify(border_08N)

border_08N_b = border_08N |> 
  sf::st_buffer(dist = 100) |> 
  rmapshaper::ms_simplify()
```

### Streams Along Border of 08N that Touches Target Watersheds

Note: these streams are quite hard to see - look for tiny red lines along the border of 08N with the other watersheds.

```{r find_streams_near_border_of_08N}
# Use the un-summarized version of the watersheds. Find the ones that 
# touch our border. Query BCDC one-by-one.
wscd_touch = wscd |> 
  sf::st_filter(border_08N_b)

wscd_all = dplyr::summarise(wscd_t)

# Cycle through the smaller watershed drainage areas that touch the border with
# 08N; get all streams that touch that portion of the border; simplify the 
# resulting tables.
if(!file.exists('data/pascale/streams_along_08N_border.gpkg') & !file.exists('../data/pascale/streams_along_08N_border.gpkg')){
  streams_along_08N_border = wscd_touch$SUB_SUB_DRAINAGE_AREA_CD |> 
    purrr::map( ~ {
      
      print(.x)
      
      this_wscd = wscd_touch[wscd_touch$SUB_SUB_DRAINAGE_AREA_CD == .x,]
      border_touching_this_wscd = sf::st_intersection(border_08N_b, this_wscd)
      
      # The first query here is quite fragmentary: it only shows the piece of 
      # each stream that touches the boundary; I'd like to get the entire stream geometry.
      streams_touching_border = bcdc_query_geodata('freshwater-atlas-stream-network') |> 
        filter(INTERSECTS(border_touching_this_wscd)) |> 
        collect() |> 
        sf::st_zm()
      
      if(nrow(streams_touching_border) > 0){
        # Simplify the results table.
        streams_touching_border = streams_touching_border |> 
          dplyr::select(GNIS_NAME, FWA_WATERSHED_CODE, WATERSHED_GROUP_ID, BLUE_LINE_KEY) |> 
          dplyr::mutate(across(-geometry, as.character)) |> 
          dplyr::mutate(wscd_name = .x)
      } else {
        streams_touching_border = NULL
      }
      return(streams_touching_border)
    })
  
  str_on_b = streams_along_08N_border |> 
    dplyr::bind_rows()
  
  if(stringr::str_detect(getwd(), 'scripts')){
    sf::write_sf(str_on_b, paste0('../data/pascale/streams_along_08N_border.gpkg'))
  } else {
    sf::write_sf(str_on_b, paste0('data/pascale/streams_along_08N_border.gpkg'))
  }
} else {
  if(file.exists('data/pascale/streams_along_08N_border.gpkg')) str_on_b = sf::read_sf('data/pascale/streams_along_08N_border.gpkg')
  if(file.exists('../data/pascale/streams_along_08N_border.gpkg')) str_on_b = sf::read_sf('../data/pascale/streams_along_08N_border.gpkg')
}

base_wd_map + 
  geom_sf(data = str_on_b, col = 'red')
```

```{r find_watershed_id_of_downstream_neighbour_of_border_streams}

# Determine which of our streams are flowing FROM 08N to any of the other watersheds.

if(!file.exists('data/pascale/border_streams_downstream_neighbour.csv') & !file.exists('../data/pascale/border_streams_downstream_neighbour.csv')){
  str_on_b_ds = str_on_b |> 
    sf::st_drop_geometry() |> 
    dplyr::select(FWA_WATERSHED_CODE) |> 
    dplyr::filter(!duplicated(FWA_WATERSHED_CODE)) |> 
    dplyr::rename(upstream_fwa_code = FWA_WATERSHED_CODE)
  
  # Join on a new field: the FWA code of the downstream stream.
  str_on_b_ds = str_on_b_ds |> 
    dplyr::left_join(
      fwa.connect::fwa_up_and_downstream_tbl
    )
  
  # Find which watershed each downstream stream is in.
  # If it's in 08N, we can drop that stream. If it's in 
  # one of our target watersheds, we need to keep it.
  str_on_b_ds$ws_of_ds_stream = NA
  
  for(i in 1:nrow(str_on_b_ds)){
    
    ds_stream = NULL
    
    print(paste0(i," of ",nrow(str_on_b_ds)))
    the_fwa_code = str_on_b_ds[i,]$downstream_fwa_code
    
    if(is.na(the_fwa_code)) next
    
    the_pattern = bcdata:::CQL(paste0("FWA_WATERSHED_CODE like '",the_fwa_code,"'"))
    
    # Get geometry of downstream stream
    ds_stream = bcdata::bcdc_query_geodata('freshwater-atlas-stream-network') |> 
      bcdata::filter(the_pattern) |>
      bcdata::collect() |> 
      sf::st_zm() |> 
      dplyr::group_by(BLUE_LINE_KEY,FWA_WATERSHED_CODE) |>
      dplyr::summarise() |> 
      dplyr::ungroup()
    
    # If our query returned multiple streams with different BLKs, some tend 
    # to be little loops that we don't need to worry about.
    # Find the longest stream and choose that one.
    ds_stream = ds_stream |> 
      dplyr::mutate(the_length = as.numeric(sf::st_length(geometry))) |> 
      dplyr::arrange(dplyr::desc(the_length)) |> 
      dplyr::slice(1)
    
    # Which watershed is/are the stream(s) in?
    ds_stream = ds_stream |> 
      sf::st_join(wscd_t)
    
    # Drop instances of streams looping back into 08N
    if(nrow(ds_stream) > 1){
      ds_stream = ds_stream |> 
        dplyr::filter(SUB_DRAINAGE_AREA_CD != '08N')
      
    }
    
    if(nrow(ds_stream) > 1){
      ds_stream = ds_stream |> 
        dplyr::group_by(BLUE_LINE_KEY, FWA_WATERSHED_CODE) |> 
        dplyr::summarise(SUB_DRAINAGE_AREA_CD = paste0(SUB_DRAINAGE_AREA_CD, collapse = ', ')) |> 
        dplyr::ungroup()
    }
    
    str_on_b_ds[i,]$ws_of_ds_stream = ds_stream$SUB_DRAINAGE_AREA_CD[1]
  }
  
  # Drop NA and 08N entries for downstream ws id
  str_on_b_ds_ws = str_on_b_ds |> 
    dplyr::filter(!is.na(ws_of_ds_stream)) |> 
    dplyr::filter(ws_of_ds_stream != '08N') |> 
    dplyr::rename(FWA_WATERSHED_CODE = upstream_fwa_code)
  
  if(stringr::str_detect(getwd(), 'scripts')){
    write.csv(str_on_b_ds_ws, paste0('../data/pascale/border_streams_downstream_neighbour.csv'), row.names = F)
  } else {
    write.csv(str_on_b_ds_ws, paste0('data/pascale/border_streams_downstream_neighbour.csv'), row.names = F)
  }
} else {
  if(file.exists('data/pascale/border_streams_downstream_neighbour.csv')) str_on_b_ds_ws = readr::read_csv('data/pascale/border_streams_downstream_neighbour.csv')
  if(file.exists('../data/pascale/border_streams_downstream_neighbour.csv')) str_on_b_ds_ws = readr::read_csv('../data/pascale/border_streams_downstream_neighbour.csv')
}
```

```{r find_fwa_codes_of_all_streams_in_graphs_that_include_border_streams}
str_on_b_outside_08N = str_on_b |> 
  dplyr::inner_join(str_on_b_ds_ws)

library(tidygraph)

all_streams_w_group_ids = tidygraph::as_tbl_graph(fwa.connect::fwa_up_and_downstream_tbl) |> 
  activate(nodes) |> 
  mutate(group_id = group_components())

all_streams_w_group_ids <- as_tibble(all_streams_w_group_ids, what = "nodes") |> 
  dplyr::rename(FWA_WATERSHED_CODE = name)

str_on_b_w_group = str_on_b_outside_08N |> 
  dplyr::left_join(all_streams_w_group_ids)

rm(all_streams_w_group_ids)

# Find the FWA_WATERSHED_CODEs of ALL streams that are present in the 
# graph components of streams whose downstream neighbour is in one of 
# our target watersheds.
codes_for_download = str_on_b_w_group |> 
  dplyr::filter(group_id %in% unique(str_on_b_w_group$group_id)) |> 
  dplyr::filter(!duplicated(FWA_WATERSHED_CODE)) |> 
  dplyr::pull(FWA_WATERSHED_CODE)
```

```{r split_watersheds_into_subwatersheds_and_then_download_all_streams}

# Split the watersheds into their subwatersheds, find intersection of those. Use those to download streams.
subws = wscd |> 
  dplyr::filter(SUB_DRAINAGE_AREA_CD %in% wscd_t$SUB_DRAINAGE_AREA_CD) |> 
  sf::st_intersection(bc)

# base_wd_map + 
#   geom_sf(data = subws, fill = 'transparent', col = 'grey')

# Try to download all streams in these watersheds. Maybe go by subwatershed so that it's not too large.
for(area_cd in unique(subws$SUB_SUB_DRAINAGE_AREA_CD)){
  
  if(!file.exists(paste0('../data/pascale/streams_in_target_watersheds/streams_',area_cd,'.gpkg')) & !file.exists(paste0('data/pascale/streams_in_target_watersheds/streams_',area_cd,'.gpkg'))){
    
    print(area_cd)
    
    the_subws = subws[subws$SUB_SUB_DRAINAGE_AREA_CD == area_cd,]
    
    ggplot() + geom_sf(data = the_subws)
    
    streams = bcdc_query_geodata('freshwater-atlas-stream-network') |> 
      bcdata::filter(INTERSECTS(the_subws)) |> 
      collect() |> 
      sf::st_zm() |> 
      dplyr::mutate(watershed_id = area_cd)
    
    # ggplot() + 
    #   geom_sf(data = the_subws, fill = 'purple', col = 'black') + geom_sf(data = streams, col='red')
    
    if(stringr::str_detect(getwd(), 'scripts')){
      sf::write_sf(streams, paste0('../data/pascale/streams_in_target_watersheds/streams_',area_cd,'.gpkg'))
    } else {
      sf::write_sf(streams, paste0('data/pascale/streams_in_target_watersheds/streams_',area_cd,'.gpkg'))
    }
  }
}
```

```{r read_in_and_join_all_streams_dropping_duplicate_geometries}

if(!file.exists('data/pascale/all_streams_in_watersheds_of_interest.gpkg') & !file.exists('../data/pascale/all_streams_in_watersheds_of_interest.gpkg')){

if(stringr::str_detect(getwd(), 'scripts')){
    stream_gpkgs = list.files('../data/pascale/streams_in_target_watersheds/',
               pattern = '.gpkg$',
               full.names = T) |> 
      lapply(sf::read_sf)
  } else {
    stream_gpkgs = list.files('data/pascale/streams_in_target_watersheds/',
               pattern = '.gpkg$',
               full.names = T) |> 
      lapply(sf::read_sf)
  }

# All streams in watersheds! Pick fewer columns and bind all rows.
all_str_in_ws = purrr::map(stream_gpkgs, ~ {
  .x |> 
    dplyr::select(GNIS_NAME, BLUE_LINE_KEY, FWA_WATERSHED_CODE, GRADIENT)
}) |> 
  dplyr::bind_rows()

# Add on graph component ids; drop components that aren't in the handful we identified as 
# border-crossing groups!
all_str_in_ws = all_str_in_ws |> 
  dplyr::left_join(all_streams_w_group_ids) |> 
  dplyr::filter(group_id %in% unique(str_on_b_w_group$group_id))

if(stringr::str_detect(getwd(), 'scripts')){
  sf::write_sf(all_str_in_ws,'../data/pascale/all_streams_in_watersheds_of_interest.gpkg')
} else {
  sf::write_sf(all_str_in_ws,'data/pascale/all_streams_in_watersheds_of_interest.gpkg')
}
} else {
  if(stringr::str_detect(getwd(), 'scripts')){
    all_str_in_ws = sf::read_sf('../data/pascale/all_streams_in_watersheds_of_interest.gpkg')
  } else {
    all_str_in_ws = sf::read_sf('data/pascale/all_streams_in_watersheds_of_interest.gpkg')
  }
}

```

```{r brainstorming}
# Things to do maybe
# 1. Find intersection points where one stream feeds into another stream. Make into points. Get elevation, 
# find directionality of streams such that higher streams flow into lower streams.

# 2. Only include named streams and see what the extent looks like?
```

### Plot of Named Streams in Target Watersheds Coloured by Group ID

```{r named_streams_plot}
if(!file.exists('data/pascale/named_streams_in_watersheds_of_interest.gpkg') & !file.exists('../data/pascale/named_streams_in_watersheds_of_interest.gpkg')){
  
  named_str = all_str_in_ws |> 
    dplyr::filter(!is.na(GNIS_NAME)) |> 
    dplyr::mutate(group_id = as.character(group_id)) |> 
    dplyr::group_by(GNIS_NAME, BLUE_LINE_KEY, FWA_WATERSHED_CODE, group_id) |> 
    dplyr::summarise()
  
  if(stringr::str_detect(getwd(), 'scripts')){
    sf::write_sf(named_str,'../data/pascale/named_streams_in_watersheds_of_interest.gpkg')
  } else {
    sf::write_sf(named_str,'data/pascale/named_streams_in_watersheds_of_interest.gpkg')
  }
} else {
  if(stringr::str_detect(getwd(), 'scripts')){
    named_str = sf::read_sf('../data/pascale/named_streams_in_watersheds_of_interest.gpkg')
  } else {
    named_str = sf::read_sf('data/pascale/named_streams_in_watersheds_of_interest.gpkg')
  }
}

# ggplot() + 
  # geom_sf(data = bc) +
  # geom_sf(data = wscd_t, aes(fill = SUB_DRAINAGE_AREA_CD), alpha = 0.35, col = 'darkgrey') +
base_wd_map +
  geom_sf(data = named_str, aes(color = group_id)) +
  geom_sf(data = named_str[named_str$GNIS_NAME %in% c("Fraser River","Columbia River"),],
          aes(color = group_id), linewidth = 1.2) +
  ggrepel::geom_label_repel(
    data = named_str[named_str$GNIS_NAME %in% c("Fraser River","Columbia River"),],
    aes(label = GNIS_NAME, geometry = geom),
    stat = "sf_coordinates",
    min.segment.length = 0,
    nudge_x = 200000, 
    nudge_y = 250000) +
  scale_fill_brewer(palette = 'Set1') + 
  ggthemes::theme_map() + 
  labs(fill = 'Watershed Code',
       color = 'Group ID',
       title = 'Connectivity of Streams in Watersheds of Interest',
       subtitle = '(only named streams displayed here for increased readability)') + 
  coord_sf(xlim = st_bbox(wscd_t)[c(1,3)],
           ylim = st_bbox(wscd_t)[c(2,4)])

```

### Tracing Course Downstream from Emerald Lake

```{r try_using_fwa_connect_trace_downstream_from_emerald_lake}

str_in_em_lake = all_str_in_ws |> 
  sf::st_filter(em_lake)

fwa_codes_in_em_lake = unique(str_in_em_lake$FWA_WATERSHED_CODE)

ds_from_em_lake = suppressMessages(fwa.connect::trace_course_downstream(fwa_code = fwa_codes_in_em_lake[2]))

base_wd_map +
  geom_sf(data = em_lake) + 
  geom_sf(data = ds_from_em_lake, aes(col = FWA_WATERSHED_CODE)) + 
  ggrepel::geom_label_repel(data = em_lake, aes(label = GNIS_NAME_1, geometry = geometry),
                            stat = 'sf_coordinates') +
  ggthemes::theme_map() + 
  theme(legend.position = 'none') + 
  labs(title = 'Tracing Course Downstream from Emerald Lake, Yoho National Park')
```

```{r clean_data}
# # Get top 50 destination waterbodies
# insp_top = insp |>
#   dplyr::filter(!GNIS_NA %in% c("Dry Storage","Pacific Ocean")) |>
#   dplyr::arrange(dplyr::desc(TotalInspections)) |>
#   dplyr::slice(1:50)
# 
# insp_top = st_transform(insp_top, 4326)
# bc = st_transform(bcmaps::bc_bound(), 4326)
# 
# insp_top = insp_top |> 
#   dplyr::mutate(colours = as.numeric(cut(TotalInspections,5)),
#                 bin_levels = as.character(cut(dig.lab = 8,TotalInspections,5))) |> 
#   dplyr::mutate(bin_levels = stringr::str_remove_all(bin_levels, '(\\(|\\])')) |> 
#   dplyr::mutate(bin_levels = stringr::str_replace(bin_levels, ',', '-')) |> 
#   dplyr::mutate(bin_levels = stringr::str_replace(bin_levels, '\\.[0-9]+', '')) |> 
#   dplyr::mutate(bin_levels = as.factor(bin_levels)) |> 
#   dplyr::mutate(bin_levels = forcats::fct_inorder(bin_levels))
```
